Constraint Programming - Inside Gecode

### 2014-05-05 paper.tex
removed C++-code and pseudo code

	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \refeq{eq:linearConstraint}. Which statements are possible to make and how handy it is to do so varies very much from the programming language in which the problem is implemented and the used toolset. Typical constraints would be
	\begin{itemize}
		\item Equalities: $x = y + 2$, $x^2 + y^2 = 1$, \dots
		\item Inequalities: $x \ne y$, \dots
		\item Relations: $x < 2 y$, $e^x < 4$, \dots
	\end{itemize}
	In \emph{Gecode} it is very easy to create for instance a nonlinear constraint based on the integer variables $a$, $b$, $c$ and $d$ \cite[p.~120]{programmingGecode}:
	\begin{lstlisting}
rel(home, a+b*(c+d) == 0);
	\end{lstlisting}
	This single statement in C++ implements the constraint
	\begin{equation}
		a + b \cdot (c + d) == 0
	\end{equation}
	which is a quite handy way to do it.

	Hence there are boolean variables available it is also possible to use logical operators to combine other constraint types to logical expressions. To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap. The related variables are defined in \reffig{fig:squares}. This problem can be expressed mathematically by \cite[p. 101]{programmingGecode}
	\begin{equation}
	\begin{split}
		x_1 + d_1 \le x_2 & \lor x_2 + d_2 \le x_1 \lor \\
		y_1 + d_1 \le y_2 & \lor y_2 + d_2 \le y_1
	\end{split}
	\end{equation}
	and in \emph{Gecode} by
	\begin{lstlisting}
rel(home,	(x1 + d1 <= x2) || 
		(x2 + d2 <= x1) || 
		(y1 + d1 <= y2) || 
		(y2 + d2 <= y1));
	\end{lstlisting}
	which is again quite obvious to a reader with some experience in a C-like programming language.

@@

	\subsubsection{Depth-First Search}
	\begin{algorithm}
		\begin{algorithmic}
			\Function{DFS}{node}
				\If{$Goal(node)$}
					\State \Return $node$
				\EndIf
				\State $s \gets Successors(node)$
				\If{$Null(s)$}
					\State \Return $nil$
				\EndIf
				\State $result \gets DFS(Left(s))$
				\If{$result \ne nil$}
					\State \Return $result$
				\Else
					\State \Return $DFS(Right(s))$
				\EndIf
			\EndFunction
		\end{algorithmic}
		\caption{Depth-first search in a binary search tree for one feasible solution, without constraint propagation}
	\end{algorithm}

@@

	\begin{algorithm}
		\begin{algorithmic}
			\Function{$LDSProbe$}{$node$, $k$}
				\If{$Goal(node)$}
				    \State \Return $node$
				\EndIf
				\State $s \gets Successors(node)$
				\If{$Null(s)$}
					\State \Return $nil$
				\EndIf
				\If{$k = 0$}
					\State \Return $LDSProbe(Left(s), 0)$
				\Else
					\State $result \gets LDSProbe(Right(s), k - 1)$
					\If{$result \ne nil$}
						\State \Return $result$
					\EndIf
					\State \Return $LDSProbe(Left(s), k)$
				\EndIf
			\EndFunction
		\end{algorithmic}
		\caption{Probe of Limited Discrepancy Search for a binary search tree from \cite{limitedDiscrepancy}, without constraint propagation}
	\end{algorithm}

	\begin{algorithm}
		\begin{algorithmic}			
			\Function{$LDS$}{$node$}
				\For{$x \gets 0\ \text{to maximum depth}$}
					\State $result \gets LDSProbe(node, x)$
					\If{$result \ne nil$}
						\State \Return $result$
					\EndIf
				\EndFor
			\EndFunction
		\end{algorithmic}
		\caption{Limited Discrepancy Search for a binary search tree from \cite{limitedDiscrepancy}, without constraint propagation}
	\end{algorithm}

>>>>

	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \refeq{eq:linearConstraint}. Which statements are possible to make and how handy it is to do so varies very much from the programming language in which the problem is implemented and the used toolset. Typical constraints would be
	\begin{itemize}
		\item Equalities: $x = y + 2$, $x^2 + y^2 = 1$, \dots
		\item Inequalities: $x \ne y$, \dots
		\item Relations: $x < 2 y$, $e^x < 4$, \dots
	\end{itemize}

	Hence there are boolean variables available it is also possible to use logical operators to combine other constraint types to logical expressions. To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap. The related variables are defined in \reffig{fig:squares}. This problem can be expressed by \cite[p. 101]{programmingGecode}
	\begin{equation}
	\begin{split}
		x_1 + d_1 \le x_2 & \lor x_2 + d_2 \le x_1 \lor \\
		y_1 + d_1 \le y_2 & \lor y_2 + d_2 \le y_1
	\end{split}
	\end{equation}

@@

@@
